### 2013-10-12 Rcode/intro.rmd
added some more examples to intro.rmd

# compare to the Bayes estimator for this problem
 mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
# note that the model parameters are in terms of the MARGINAL not of the prior
 trueEff = effectSize(s,mTrue)[,1]
 trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
 par(mfrow = c(2,1))
 plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
 lines(s, effCurve, col = 2)
 plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))
 lines(s, fdrCurve, col = 2)

@@

Don't need to worry about overfitting - only need to worry
about whether model is "complex enough".

Because of the uni-modal constraint, the number of mixture components
is generally not critical, at least provided it is ``sufficiently large". 
Indeed, as the number of components tends to infinity, the likelihood
is bounded above, and the procedure will converge to a sensible limit.
The following example illustrates this.
We take 100 observations from 0.5N(-1,1)+0.5N(1,1)
and add N(0,1) noise.
```{r}
sampsize=500
beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
betahat = beta+rnorm(2*sampsize)

beta.ash = ash(betahat,1)
beta.ash.u=list()
for(i in 1:10){
  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
}

@@

@@

beta.ash = ash(betahat,1)
beta.ash.u=list()
for(i in 1:10){
  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",prior="uniform",sigmaavec = seq(0.1,5,length=10*i))
}

@@

beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
beta.ash.vb.null = ash(ss$betahat, ss$betasd,VB=TRUE, prior=NULL)

>>>>

# compare to the Bayes estimator for this problem
 mTrue = list(pi = c(10,1)/11, mu = c(0,3), sig = c(1,1))
# note that the model parameters are in terms of the MARGINAL not of the prior
 trueEff = effectSize(s,mTrue)[,1]
 trueFdr = fdrMixModel(s,mTrue,c(TRUE,FALSE))
 par(mfrow = c(2,1))
 plot(s, trueEff, t = 'l', main = "Effect Size - Black is true", xlab = "z", ylab = "E(delta|z)", ylim = c(-3,3))
 lines(s, effCurve, col = 2)
 plot(s, trueFdr, t = 'l', main = "fdr - Black is true", xlab = "z", ylab = "fdr", ylim = c(0,1))

@@

Because of the uni-modal constraint, the number of mixture components
is generally not critical, at least provided it is ``sufficiently large". 
Indeed, as the number of components tends to infinity, the likelihood
is bounded above, and even for large numbers of components and small amounts of data the inferred underlying distributions tend not to be too crazy, showing few signs of ``overfitting".

However, it is true that the normal distributions tend to yield smoother estimated underlying distributions. Similarly, using the posterior mean for $\pi$, rather
than the maximum likelihood estimate, tends to lead to somewhat smoother fitted distributions. In addition, with small amounts of data the underlying
distributions are inevitably not well determined by the data, and the fits may vary depending on the underlying assumptions made (e.g. number of components or distribution of the components); however, even then shrinkage-based estimates
of the betas can be relatively robust.

The following example illustrates this.
We take 100 observations from 0.5N(-1,1)+0.5N(1,1)
and add N(0,1) noise.
```{r}
set.seed(200)
sampsize=50
beta = c(rnorm(sampsize,-1,1),rnorm(sampsize,1,1))
betahat = beta+rnorm(2*sampsize)

beta.ash.default = ash(betahat,1) #run with default parameters for comparison
beta.ash.vb.default = ash(betahat,1,VB=TRUE) #run with default parameters for comparison

beta.ash.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100))
beta.ash.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100))
beta.ash.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100))

beta.ash.vb.100 = ash(betahat,1,mixsd = seq(0.1,5,length=100),VB=TRUE)
beta.ash.vb.100u = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=100),VB=TRUE)
beta.ash.vb.100hu = ash(betahat,1,mixcompdist="halfuniform",mixsd = seq(0.1,5,length=100),VB=TRUE)

x=seq(-4,4,length=1000)
plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities vs truth",type="l",ylab="density")
lines(x,density(beta.ash.default,x),col=2)
lines(x,density(beta.ash.100,x),col=2)
lines(x,density(beta.ash.100u,x),col=3)
lines(x,density(beta.ash.100hu,x),col=4)

plot(x,0.5*dnorm(x,-1,1)+0.5*dnorm(x,1,1),col=1,ylim=c(0,0.45),main="Fitted densities (VB) vs truth",type="l",ylab="density")
lines(x,density(beta.ash.vb.default,x),col=2)
lines(x,density(beta.ash.vb.100,x),col=2,lty=2)
lines(x,density(beta.ash.vb.100u,x),col=3,lty=2)
lines(x,density(beta.ash.vb.100hu,x),col=4,lty=2)

plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs vs truth",type="l",ylab="density")
lines(x,cdf.ash(beta.ash.100,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.100u,x),col=3,lty=2)
lines(x,cdf.ash(beta.ash.100hu,x),col=4,lty=2)

plot(x,0.5*pnorm(x,-1,1)+0.5*pnorm(x,1,1),col=1,main="Fitted cdfs (VB) vs truth",type="l",ylab="density")
lines(x,cdf.ash(beta.ash.vb.default,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.vb.100,x),col=2,lty=2)
lines(x,cdf.ash(beta.ash.vb.100u,x),col=3,lty=2)
lines(x,cdf.ash(beta.ash.vb.100hu,x),col=4,lty=2)

c(loglik.ash(beta.ash.100,betahat,1),get_loglik(beta.ash.100u),get_loglik(beta.ash.100hu))
c(loglik.ash(beta.ash.vb.100,betahat,1),loglik.ash(beta.ash.vb.100u,betahat,1),loglik.ash(beta.ash.vb.100hu,betahat,1))

rmse=function(a,b){sqrt(mean((a$PosteriorMean-b)^2))}
c(rmse(beta.ash.default,beta),rmse(beta.ash.100,beta),rmse(beta.ash.100u,beta),rmse(beta.ash.100hu,beta),rmse(beta.ash.vb.default,beta),rmse(beta.ash.vb.100,beta),rmse(beta.ash.vb.100u,beta),rmse(beta.ash.vb.100hu,beta))

plot(beta.ash.default$PosteriorMean,beta.ash.100u$PosteriorMean,col=3,cex=0.4,xlim=c(-4,4),ylim=c(-4,4))
points(beta.ash.default$PosteriorMean,beta.ash.100hu$PosteriorMean,col=4,cex=0.4)
abline(a=0,b=1)

plot(betahat,ifelse(pnorm(betahat)<0.5,pnorm(betahat),1-pnorm(betahat)))
points(betahat,beta.ash.default$localfsr,col=2)
points(betahat,beta.ash.100u$localfsr,col=3)
points(betahat,beta.ash.100hu$localfsr,col=4)

@@

so it seems that the localfsr is almost the same as for a "flat" prior $g$.
That is, in this case, the underlying distribution is flat relative to the
error terms, and so there is little shrinkage effect except at the tails
(where the fsr is basically 0 anyway).

@@

beta.ash = ash(betahat,1)
beta.ash.u=list()
for(i in 1:10){
  beta.ash.u[[i]] = ash(betahat,1,mixcompdist="uniform",mixsd = seq(0.1,5,length=10*i))
}

@@

beta.ash=ash(ss$betahat,ss$betasd,auto=FALSE)
beta.ash.auto = ash(ss$betahat, ss$betasd,auto=TRUE)
#these to test the VB version 
beta.ash.vb.uniform = ash(ss$betahat, ss$betasd,VB=TRUE, prior="uniform")
beta.ash.vb.nullbiased = ash(ss$betahat, ss$betasd,VB=TRUE, prior="nullbiased")
